MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.45\textwidth]{images_pvis/workflow.png}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Fig.~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
 In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$. In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Fig. \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue. Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=0.3\textwidth]{imagesMT2014/image4}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Fig..~\ref{fig:comparison_orient} shows the MetaTracts in Fig..~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step. We set ``k" to be  10, 15, 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[hb]
	\centering
	\includegraphics[width=0.45\textwidth,  trim = 0mm 10mm 0mm 10mm, clip]{images_pvis/comparison_proximity.pdf}
	\caption{MetaTracts data (left) and its straight-forward hierarchical clustering for various numbers of clusters.}
	\label{fig:comparison}
\end{figure} 
\begin{figure}[hb]
	\centering
	\includegraphics[width=0.45\textwidth]{images_pvis/comparison_orient.pdf}
	\caption{Direct K-means clustering of MetaTracts (left) embedded in a m dimensional space using laplacian eigenmaps by proximity into 10,15 and 20 clusters.}
	\label{fig:comparison_orient}
\end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=0.45\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

>>>>

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Fig.~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
 In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$.The procedure trivially extends to $\mathbb{R}^3$. In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Fig. \ref{fig:algo}, all regions which have unreliable Hessian have been marked in blue. Fig.~\ref{fig:algo}c shows a single cylinder of a MetaTracts. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$, the green dashed rectangle shows the rectangle generated at $C_p$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Sec.~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
In comparison Fig.~\ref{fig:comparison} shows the MetaTracts in Fig.~\ref{fig:length_distribution}a being hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Fig..~\ref{fig:comparison_orient} shows the MetaTracts in Fig..~\ref{fig:length_distribution}a, being clustered using K-means clustering after being embedded  in a m-dimensional (lower) space using proximity alone, without a second hierarchical clustering step. We set ``k" to be  10, 15, 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[hb]
	\centering
	\includegraphics[width=0.45\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
	\label{fig:comparison}
\end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=0.45\linewidth]{images_pvis/dataset2.pdf}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}
